---
title: "P8105 HW2 hl3439"
author: Hanyu Lu
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset. 

```{r}
trashwheel_df = 
	read_xlsx(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)
```

Read precipitation data! For 2018 and 2017. 

```{r}
precip_2018 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)
precip_2017 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

Now combine annual precipitation.

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)
precip_df = 
	bind_rows(precip_2018, precip_2017)
left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. 

## Problem 2

```{r}
ny_subway_df = 
  read.csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(-division, -staffing, -staff_hours, -entrance_longitude, -entrance_latitude, -entrance_location, -station_location, -corner, -east_west_street, -north_south_street, -free_crossover, -ada_notes, -exit_only) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE)) 
  
```

This dataset contains the following variables `r names(ny_subway_df)`.  After importing the dataset, I clean names to make column names be of the same format and remove columns irrelevant to the problem.  Additionally, I convert the entry variable from character to a logical variable.  The dimension of the resulting data has `r nrow(ny_subway_df)` rows and `r ncol(ny_subway_df)` columns.  These data are not yet super-tidy because there are 11 routes columns which make it hard for humans to read.  These data will be tidied further later.

* There are `r nrow(distinct(ny_subway_df, line, station_name))` distinct stations.  

```{r}
distinct_station = 
  distinct(ny_subway_df, line, station_name, ada, .keep_all = TRUE)
```

* `r sum(pull(distinct_station, ada))` stations are ADA compliant.

```{r}
ny_subway_df =
  mutate(ny_subway_df, vending = recode(vending, "YES" = TRUE, "NO" = FALSE))

no_vending_entrance =
  filter(ny_subway_df, vending == FALSE)
no_vending_num = nrow(no_vending_entrance)

allow_entrance =
  sum(pull(no_vending_entrance, entry))
```

* `r allow_entrance/no_vending_num` of station entrances / exits without vending allow entrance.

```{r}
ny_subway_df =
  mutate(
    ny_subway_df, 
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)
    ) %>% 
  mutate_if(is.character, list(~na_if(.,""))) 

ny_subway_tidy = 
  pivot_longer(
    ny_subway_df,
    route1:route11,
    names_to = "route_number",
    values_to = "route_name",
  )


```

The above code chunk reformats data so that route number and route name are distinct variables.

```{r}
distinct_station_tidy =
  distinct(ny_subway_tidy, line, station_name, ada, route_name, .keep_all = TRUE)

A_station = 
  filter(distinct_station_tidy, route_name == "A")
A_station_num = nrow(A_station)

A_station_ada =
  filter(distinct_station_tidy, route_name == "A", ada == TRUE)
A_station_ada_num = nrow(A_station_ada)
```

There are `r A_station_num` distinct stations serve the A train.  Of the stations that serve the A train, `r A_station_ada_num` are ADA compliant

## Problem 3
```{r}
pols_month_df = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, sep="-", into = c("year", "month", "day")) %>% 
  mutate(
    month = month.abb[as.factor(month)],
    president = case_when(prez_gop == "1" ~ "gop", prez_gop == "0" ~ "dem")
  )  %>% 
  select(-prez_gop, -prez_dem, -day) 

snp_df = read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, sep="/", into = c("month", "day", "year")) %>% 
  mutate(
    month = month.abb[as.factor(month)],
   ) %>% 
  arrange(year, month) %>% 
  relocate(year, month)
  

unemployment_df = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment"
  )

```

```{r}

```

